---
title: 'Class 9: Unsupervised Learning II'
author: "Quoc Tran"
date: "April 30, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Preparing the data
```{r}
wisc.df <- read.csv("WisconsinCancer.csv")

wisc.data <- as.matrix(wisc.df[,c(3:32)], nrow=nrow(wisc.df), ncol=30)
rownames(wisc.data) <- wisc.df$id

diagnosis <- wisc.df$diagnosis
diagnosis <- as.numeric(diagnosis=="M")
```

##Exploratory data analysis
Q1. How many observations are in this dataset?
```{r}
nrow(wisc.data)
```

Q2. How many variables/features in the data are suffixed with _mean?
```{r}
length(grep("_mean", colnames(wisc.data)))
```

Q3. How many of the observations have a malignant diagnosis?
```{r}
sum(diagnosis)
```

##Performing PCA
```{r}
wisc.pr <- prcomp(wisc.data, scale.=TRUE)
summary(wisc.pr)
```

Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

0.4427

Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

3 principal components (0.72636)

Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

7 principal components (0.91010)

##Interpreting PCA Results
```{r}
biplot(wisc.pr)
```

Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?

It's very messy and very hard to understand because all of the principal components and all of the data points are being plotted.

**PC1 vs PC2 (red = malignant)**
```{r}
plot(wisc.pr$x[,1:2], col=(diagnosis+1), xlab="PC1", ylab="PC2")
```

**PC1 vs PC3 (red = malignant)**
```{r}
plot(wisc.pr$x[,c(1,3)], col=(diagnosis+1), xlab="PC1", ylab="PC3")
```

Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?

There is a clear delineation between malignant and benign tumors along principal component 1.

##Variance explained

```{r}
pr.var <- wisc.pr$sdev^2 # standard deviation squared = variation
pve <- pr.var/sum(pr.var) # proportion of variance
plot(pve, xlab="Principal Component", ylab="Proportion of Variance Explained", ylim=c(0,1), typ="o", main="Scree Plot")
```

```{r}
plot(cumsum(pve), xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained", ylim=c(0,1), typ="o", main="Scree Plot")
```

##Communicating PCA Results

Q9. For the first principal component, what is the component of the loading vector (i.e.  wisc.pr$rotation[,1]) for the feature concave.points_mean?
```{r}
wisc.pr$rotation[,(colnames(wisc.data) == "concave.points_mean")]
```


Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?
```{r}
pc.min <- 0
index <- 1
while (pc.min < 0.8) {
  pc.min <- pc.min + pve[index]
  index = index + 1
}
paste(index-1, "principal components are required to explain", round(pc.min*100, 1), "% of the data.")
```

##Hierarchical clustering of case data

```{r}
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method="complete")
```

Q11. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?
```{r}
plot(wisc.hclust)
abline(h=19, col="red")
```

##Selecting number of clusters

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)
table(wisc.hclust.clusters, diagnosis)
```

Q12. Can you find a better cluster vs diagnoses match with by cutting into a different number of clusters between 2 and 10?
```{r}
for (i in c(2:10)) {
  wisc.hclust.clusters <- cutree(wisc.hclust, k=i)
  table(wisc.hclust.clusters, diagnosis)
}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4) # back to normal
```

##K-means clustering and comparing results

```{r}
wisc.km <- kmeans(scale(wisc.data), centers=2, nstart=20)
table(wisc.km$cluster, diagnosis)
table(wisc.km$cluster, wisc.hclust.clusters)
```

##Clustering on PCA results

```{r}
pc.min <- 0
index <- 1
while (pc.min < 0.9) {
  pc.min <- pc.min + pve[index]
  index = index + 1
}
paste(index-1, "principal components are required to explain", round(pc.min*100, 1), "% of the data.")
```
```{r}
data.dist <- dist(wisc.pr$x[,1:7])
wisc.pr.hclust <- hclust(data.dist, method="ward.D2")
plot(wisc.pr.hclust)
```

**PC1 vs PC2 (red = group2**
```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
table(grps, diagnosis)
plot(wisc.pr$x[,1:2], col=grps)
```

## Sensitivity/Specificity

Q16. Which of your analysis procedures resulted in a clustering model with the best specificity? How about sensitivity?

Best specificity: hclust

Best sensitivity: kmeans

## Prediction
```{r}
new <- read.csv("new_samples.csv")
npc <- predict(wisc.pr, newdata=new)
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16, main="blue = new patients")
```

Q17. Which of these new patients should we prioritize for follow up based on your results?

We should prioritize Patient 1 (rightmost blue point) for follow up.
