---
title: 'Class 12: Structural Bioinformatics II'
author: "Quoc Tran"
date: "5/9/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading 1HSG (HIV-1 protease + indinavir) from PDB
```{r Load 1HSG}
library(bio3d)
file.name <- get.pdb("1HSG")
hiv <- read.pdb(file.name)
summary(hiv)
```

**Q1: What is the name of the two non protein resid values in this structure? What does resid
correspond to and how would you get a listing of all residue values in this structure?**

"HOH" and "MK1". "resid" corresponds to the amino acid residue that an atom belongs to.

Trimming PDB files to seperate protein and ligand PDB files
```{r Prepare Protein and Ligand PDB Files}
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")
write.pdb(prot, file="1hsg_protein.pdb")
write.pdb(lig, file="1hsg_ligand.pdb")
```

all.pdbqt comes from AutoDock Vina (using config.txt)
```{r Inspecting Docking Results}
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

Finding root mean square distance for each ligand in all.pdbqt vs the original ligand in 1HSG
```{r Finding Root Mean Squared}
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```
These are RMSD for each ligand in all.pdbqt vs the original ligan in 1HSG!
