---
title: 'Class 7: R Functions and Packages'
author: "Quoc Tran"
date: "April 23, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functions revisited

We will source a file online with our functions from last time (downloaded the R file)

```{r source rescale.R}
source("rescale.R")
```

Try out the rescale() function from last time

```{r use rescale()}
rescale(1:10)
```

rescale2() catches any strings inside input arguments before running bulk of function

```{r use rescale2()}
rescale2("hi")
```

# Find missing NA values in two vectors

Start with a simple example of the larger problem

```{r simple example}
x <- c(1,2,NA,3,NA)
y <- c(NA,3,NA,3,4)

is.na(x)
is.na(y) # should return TRUE FALSE TRUE FALSE FALSE
```

Use the AND operator to find when both are true

```{r use AND operator}
is.na(x) & is.na(y) # expect: F F T F F
```

Use sum() to find total number of NA's in both

```{r use sum()}
sum(is.na(x) & is.na(y))
```

Now we can make this into a function

```{r first function}
  both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
  }
```

```{r test the function}
x <- c(NA, 1, 2, 3, NA)
y <- c(NA, 1, NA, 2, NA)

both_na(x, y) # expect: 2
```

```{r test on various inputs}
x <- c(NA, NA, NA)
y <- c(1, NA, NA, NA)

both_na(x, y) # expect: 2



both_na(x, y) # expect: 2
```

The function recycles x to match the length of y! 1st index becomes the first missing one, 2nd becomes one after that

We need to stop it if this happens

```{r fixed function}
x <- c(NA, NA, NA)
y <- c(1, NA, NA, NA, NA, NA)
both_na2(x,y)
```

both_na3 gives a better output with which()

```{r fixed function with better ouput}
x <- c(NA, NA, NA)
y <- c(1, NA, NA)
both_na3(x,y)
```

# Find intersections of two data sets

```{r test data}
df1
df2
```

Get an even simpler data set

```{r simpler test data}
x <- df1$IDs
y <- df2$IDs
```

Search for existing functionality: intersect()

```{r use intersect()}
intersect(x,y)
```

For our real function we want more data, expression in different experiements, etc.

We need the indices of the vector where the intersection happens: use %in%

```{r using %in%}
x %in% y
x[x %in% y]
```

Use cbind() and Rstudio shortcut Code > Extract Function

```{r making gene_intersect()}
gene_intersect <- function(x, y) {
  cbind(x[x %in% y],
        y[y %in% x])
}

gene_intersect(df1$IDs, df2$IDs)
```

gene_interesect2() is polished

```{r gene_intersect2()}
gene_intersect2(df1,df2)
```

gene_interesect3() and gene_intersect4() exist too

# Calculate grade after dropping lowest score
```{r test data}
student1 <- c(100,100,100,100,100,100,100,90)
student2 <- c(100,90,90,90,90,90,97,80)
```

Use min() to find lowest value

```{r using min()}
(sum(student1) - min(student1))/(length(student1) - 1)
```

